\section{Compilation}

This section deals with the more holistic side of the compilation process,
which transforms grammar into assembly.

\subsection{Named or Anonymous Grammars}

gpeg allows two types of grammars:

\begin{itemize}
\item Named grammars, which consist of a list of named rules.
\item Anonymous grammars, which consist of a single expression.
\end{itemize}

\subsection{Imports}

Imports are treated more or less in the way that C treats '\#include's:
file contents are substituted in place, and re-evaluated. The caveat
is that this must happen at rule definition level: you cannot have
substitutions of file contents, for example, in a sequence of matchers,
or as part of a matcher (for example, inside a string).

Other caveats to the import function are: a limit to the amount of
recursion allowed, and all importing happens in the same namespace
(therefore, any recursion should result in a namespace clash error
anyway).

\subsection{Special Keywords}

\subsubsection{The '\_\_prefix' Rule}

The \_\_prefix rule will define a pattern that is called before
the evaluation of every rule that follows it. This is typically
done to remove things like whitespace and comments while parsing
things like source code or text based data formats (XML, JSON, etc).

\subsubsection{The '\_\_main' Rule}

The '\_\_main' rule, which does not have to be defined, but if it is,
it will be called first. This may be convenient when the \_\_prefix and
first rule to be called cause a grammar rule topological conflict.

\subsubsection{The '\_\_end' Matcher}

The '\_\_end' matcher, which is not a matcher but actually more like
inline assembly, is nevertheless defined part of a sequence
of matchers. It successfully ends execution of the bytecode.
The compiler simply inserts an 'end' assembly instruction at that point.

\subsection{Determining the First Rule}

A named grammar will call as the first rule:

\begin{itemize}
\item The rule called '\_\_main', or in absence of such a rule:
\item The first defined rule.
\end{itemize}

The compiler will emit a 'call' instruction to the first rule as the
first instruction, followed by an 'end' instruction ('end 0').
